### 2025-03-21 thesis/main.tex
thesis updates

 %! TeX program = xelatex

\DocumentMetadata{lang=en-CA}

\documentclass[12pt,oneside,letterpaper]{report}

\input{formatting}
\input{acronyms}
\usepackage[single]{accents}

% useful commands
% below are for statement
\newcommand{\schwa}{ə}
\newcommand{\eng}{\ng{}}
%\newcommand{\overcomma}[1]{${\rm \accentset{\mbox{\normalfont ,}}{#1}}$}
\newcommand{\Lekwnen}{L\schwa{}\overcomma{k}\textsuperscript{w}\schwa{}\ng{}\schwa{}n}
\newcommand{\skipline}{\vspace{\baselineskip}}

\def\overcomma#1{%
  \setbox0=\hbox{#1}% Save the base letter in box0.
  \dimen0=\wd0       % Get its width.
  \vbox{%
    \hbox to \dimen0{\hfil,\hfil}% Create an hbox of the same width with a centered comma.
    \nointerlineskip            % Prevent extra vertical space.
    \box0                       % Place the base letter.
  }%
}

% for pandoc
\newcommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\newcommand{\about}[1]{${\sim} #1$}

\title{From tides to tails: \\
    Does the Milky Way affect Sculptor and Ursa Minor?
}
\author{Daniel A. Boyea}

\date{\today}

\begin{document}
% \linespread{1.25}

\pagenumbering{roman}

\makeatletter
\begin{titlepage}
   \begin{center}
       \textbf{\large \@title}\\
        \skipline

        by\\
        \skipline

       \@author\\
       B.Sc. The Ohio State University, 2023\\
       \vspace*{3\baselineskip}
    A Thesis Submitted in Partial Fulfillment of the Requirements for the Degree of\\
    \skipline

    MASTER OF SCIENCE \\
    \skipline

    in the Department of Physics and Astronomy\\
       \vfill
       {\small
       ©\@author, 2025\\
       University of Victoria\\
   }
   \skipline
       {\small
       All rights reserved. This thesis may not be reproduced in whole or in part,
   by photocopy or other means, without the permission of the author.} \\
\skipline
   \end{center}
We acknowledge and respect the \Lekwnen{} (Songhees and X\textsuperscript{w}seps\schwa{}m/
Esquimalt) Peoples on whose territory the university stands, and the
\Lekwnen{} and ${\underaccent{\bar}{\rm W}}$S\'ANE\'C Peoples whose historical relationships with the
land continue to this day. 
\end{titlepage}

\addtocounter{page}{1}

\addcontentsline{toc}{chapter}{Supervisory Committee}
\begin{centering}
\textbf{\@title}
\skipline
by\\
\skipline
\@author\\
B.Sc. The Ohio State University, 2023\\
\vspace*{3\baselineskip}
\end{centering}

\subsection*{Supervisory Committee}

Dr. Julio Navarro, Supervisor\\
Department of Astronomy \\
\skipline

\noindent Dr. Kim Venn, Co-Supervisor\\
Department of Astronomy

% Abstract of the paper
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
% context
\input{abstract.tex}

% \chapter*{Acknowledgements}
% \addcontentsline{toc}{chapter}{Acknowledgements}
%

%% Lists

\tableofcontents
\listoffigures
\listoftables

%\chapter*{Symbols}
%
%
%\setlength{\tabcolsep}{0pt}
%\begin{longtable}{p{0.2\textwidth} p{0.8\textwidth}}
%$\rho$ & 3-dimensional mass density. \\
%
%$\Sigma$ & 2-dimensional density \\
%\end{longtable}
%
%\setlength{\tabcolsep}{6pt}
%\renewcommand*{\arraystretch}{1}
%
%
%\printglossary[type=\acronymtype,nonumberlist]
%\printglossary
%
%
\newpage

\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\chapter{Introduction}
\input{introduction}

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

\newpage
\bibliographystyle{aasjournal}
\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{main}

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix
% \chapter*{Appendix}
% \addcontentsline{toc}{chapter}{Appendix}
% \renewcommand{\thesection}{A.\arabic{section}}
% \renewcommand\thefigure{A.\arabic{figure}}    
% \renewcommand\theequation{A.\arabic{equation}}    
% \setcounter{figure}{0}
% \setcounter{equation}{0}
% 
% \renewcommand*{\theHsection}{A.\arabic{section}}%
% \renewcommand*{\theHsubsection}{A.\arabic{subsection}}%
% \renewcommand*{\theHfigure}{A.\arabic{figure}}%
% \renewcommand*{\theHequation}{A.\arabic{equation}}%

\chapter{Model parameters}\label{sec:jack}

\chapter{Font properties}

textwidth: \the\textwidth

\newlength{\myw}
\newlength{\myl}

---\rule[2.4pt]{12pt}{0.6pt}
\settowidth{\myw}{---}\the\myw

{\bf ---}\rule[2.2pt]{12pt}{1.1pt}

\settoheight{\myl}{---}\the\myl

-\rule[2.4pt]{3pt}{0.75pt}
I\rule{1.2pt}{8pt}
{\bf I}\rule{2.0pt}{8pt}
|\rule[-2pt]{0.6pt}{12pt}

⁄\end{document}