### 2025-05-31 thesis/layout.tex
thesis updates / reorganization of RV modeling

 %! TeX program = xelatex

\DocumentMetadata{lang=en-CA}

\documentclass[12pt,oneside,letterpaper]{report}

\input{formatting}
\input{acronyms}
\usepackage[single]{accents}

% useful commands
\usepackage{geometry}
\newgeometry{noheadfoot=true}
% below are for statement
\newcommand{\schwa}{ə}
\newcommand{\eng}{\ng{}}
%\newcommand{\overcomma}[1]{${\rm \accentset{\mbox{\normalfont ,}}{#1}}$}
\newcommand{\Lekwnen}{L\schwa{}\overcomma{k}\textsuperscript{w}\schwa{}\ng{}\schwa{}n}
\newcommand{\skipline}{\vspace{\baselineskip}}

\usepackage{layout}

\def\overcomma#1{%
  \setbox0=\hbox{#1}% Save the base letter in box0.
  \dimen0=\wd0       % Get its width.
  \vbox{%
    \hbox to \dimen0{\hfil,\hfil}% Create an hbox of the same width with a centered comma.
    \nointerlineskip            % Prevent extra vertical space.
    \box0                       % Place the base letter.
  }%
}

\title{Timid tides or tall tails? \\
    Investigating the outer reaches of the Sculptor and Ursa Minor dwarf galaxies
}
\author{Daniel A. Boyea}

\date{\today}

\begin{document}
\pagestyle{plain}
% \linespread{1.25}

\pagenumbering{roman}

\makeatletter
\begin{titlepage}
   \begin{center}
       \textbf{\large \@title}\\
        \skipline

        by\\
        \skipline

       \@author\\
       B.Sc. The Ohio State University, 2023\\
       \vspace*{3\baselineskip}
    A Thesis Submitted in Partial Fulfillment of the Requirements for the Degree of\\
    \skipline

    MASTER OF SCIENCE \\
    \skipline

    in the Department of Physics and Astronomy\\
       \vfill
       {\small
       ©\@author, 2025\\
       University of Victoria\\
   }
   \skipline
       {\small
       All rights reserved. This thesis may not be reproduced in whole or in part,
   by photocopy or other means, without the permission of the author.} \\
\skipline
   \end{center}
We acknowledge and respect the \Lekwnen{} (Songhees and X\textsuperscript{w}seps\schwa{}m/
Esquimalt) Peoples on whose territory the university stands, and the
\Lekwnen{} and ${\underaccent{\bar}{\rm W}}$S\'ANE\'C Peoples whose historical relationships with the
land continue to this day. 
\end{titlepage}

\addtocounter{page}{1}

\addcontentsline{toc}{chapter}{Supervisory Committee}
\begin{centering}
\textbf{\@title}\\
\skipline
by\\
\skipline
\@author\\
B.Sc. The Ohio State University, 2023\\
\vspace*{3\baselineskip}
\end{centering}

\subsection*{Supervisory Committee}
\skipline

Dr. Julio Navarro, Supervisor\\
Department of Astronomy \\
\skipline

\noindent Dr. Kim Venn, Co-Supervisor\\
Department of Astronomy\\

\skipline

\noindent External Examiner \\
Outside Department

% Abstract of the paper
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
% context

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

%Julio Navarro, my supervisor. Rapha and other collegues

%Friends and family and collegues (fill in later).

This work has made use of data from the European Space Agency (ESA) mission
{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
has been provided by national institutions, in particular the institutions
participating in the {\it Gaia} Multilateral Agreement. 
\citep{gaiacollaboration+2016, gaiacollaboration+2023}.

This research made use of hips2fits,\footnote{https://alasky.cds.unistra.fr/hips-image-services/hips2fits} a service provided by CDS.

%% Lists

%\chapter*{Symbols}
%
%
%\setlength{\tabcolsep}{0pt}
%\begin{longtable}{p{0.2\textwidth} p{0.8\textwidth}}
%$\rho$ & 3-dimensional mass density. \\
%
%$\Sigma$ & 2-dimensional density \\
%\end{longtable}
%
%\setlength{\tabcolsep}{6pt}
%\renewcommand*{\arraystretch}{1}
%
%
%\printglossary[type=\acronymtype,nonumberlist]
%\printglossary
%
%
\newpage

\pagenumbering{arabic}

\layout

⁄\end{document}