### 2025-09-02 thesis/abstract.tex
julio sep 1 edits to §1, 2

The satellite galaxies of the Milky Way (MW), with the exception of the
Magellanic Clouds, are all dwarf spheroidals (dSphs)---gas-free,
non-rotating stellar systems whose gravity is dominated by dark matter.
The light profiles of dSphs are usually well-approximated by an
exponential law, with a sharp outer cutoff. Yet systems like the
Sculptor (Scl) and Ursa Minor (UMi) dSphs have been found to host member
stars far from the main body, out to \(\sim10\) effective radii,
indicative of outer deviations from an exponential law. The origin of
these extended profiles is not clear, and may reflect either the
influence of Galactic tides or, instead, an innate feature. In this
thesis, we review the existence of extended light profiles in Scl and
UMi, validating the Bayesian membership catalogue of
\citet{jensen+2024}. To evaluate if tides indeed can explain these
density features, we conduct idealized N-body simulation of both
galaxies in the tidal field of the Milky Way. We find that neither dwarf
has been subjected to tidal forces strong enough to create the observed
extended light profiles from initially exponential profiles. One
complication is that Sculptor's orbit is strongly influenced by the
presence of the Large Magellanic Cloud (LMC). Because the LMC's mass is
poorly constrained, Sculptor's long term orbital history is highly
uncertain. Despite this uncertainty, our models still suggest that the
combined tides of the LMC and MW are unable to explain Scl's outer
profile. We conclude that the extended light profiles of Scl and UMi are
innate, likely reflecting their assembly histories. Mergers, accretion
events, or complex, multi-component star formation episodes are possible
explanations for the observations ed extended light profiles.