### 2025-09-27 thesis/abstract.tex
a compltet thesis draft

The satellite galaxies of the Milky Way (MW), except the Magellanic
Clouds, are all dwarf spheroidals (dSphs)---gas-free, non-rotating,
dark-matter dominated stellar systems. The light profiles of dSphs
typically follow an exponential law, with the outer density sharply
declining . Yet systems like the Sculptor (Scl) and Ursa Minor (UMi)
dSphs host member stars out to \(\sim10\) effective radii, indicating
outer deviations from an exponential law. The origin of these extended
profiles is unclear, possibly arising from Galactic tides or intrinsic
properties. In this thesis, we review the evidence for extended light
profiles in Scl and UMi, validating the Bayesian membership catalogue of
\citet{jensen+2024}. To evaluate if tides can produce these density
features, we conduct idealized N-body simulation of both galaxies in the
tidal field of the Milky Way. We find that neither dwarf experiences
tides sufficient to affect their stellar distribution. One complication
is that Sculptor's orbit is strongly influenced by the presence of the
Large Magellanic Cloud (LMC). Our models still suggest that the combined
tides of the LMC and MW are unable to explain Scl's outer profile. We
conclude that the extended light profiles of Scl and UMi not of tidal
origin. They are instead likely innate, possibly explained by past
mergers, accretion events, or episodic star formation.