### 2025-09-29 thesis/results.tex

As discussed in Chapters \ref{sec:introduction}, \ref{sec:observations},
we test whether Galactic tides are responsible for the extended density
profiles of Scl and UMi. In this Chapter, we analyze tailored N-body
simulations, using the methods described in Chapter \ref{sec:methods},
to assess the tidal impact of the Galactic potential. To anticipate our
main conclusion, we find that tides drive dark matter loss in both
systems but leave their compact stellar components largely unaffected.
The Large Magellanic Cloud (LMC) perturbs the orbit of Scl, and somewhat
UMi, yet the resulting tidal effects are weaker than in a Milky Way-only
potential. Our simulations demonstrate that recent tides are unlikely to
have altered the stellar structure of Scl or UMi.

In this Chapter, we consider Scl first, describing tidal effects from
the MW on dark matter and stellar components, and then consider the
effects from the LMC. We then analyze UMi in a similar manner,
considering in turn the dark matter evolution, stellar evolution, and
orbital effects of the LMC.

\section{Tidal effects on Sculptor}\label{tidal-effects-on-sculptor}

\subsection{Evolution of Sculptor's dark matter
halo}\label{evolution-of-sculptors-dark-matter-halo}

As a representation of an extreme tidal history, we initially
investigate the \smallperi{} Scl model.

Scl experiences moderate tidal mass loss after 10 Gyrs of evolution.
Fig.~\ref{fig:scl_sim_images} shows the stripping of dark matter and
formation of diffuse streams trailing and leading Sculptor's orbit.
Because tidal stripping can be described as a gradual removal of the
least bound particles, most mass loss occurs in the outer halo.

N-body models may deviate from a point-particle trajectory due to
dynamical self-friction \citep[e.g.,][]{white1983, miller+2020}.
However, this effect is slight for Scl, which ends near the observed
position, without adjusting the initial conditions (the green point in
Fig.~\ref{fig:scl_sim_images}).

The inner density cusp is tidally resilient.
Fig.~\ref{fig:scl_tidal_track} shows the initial and final circular
velocity profiles, and the evolution of the maximum circular velocity.
The maximum velocity drops from \(31\,\kms\) to \(22\,\kms\), evolving
along the tidal track from \citet{EN2021}. The final circular velocity
profile resembles the initial with an inner cusp, but has a sharper
outer truncation. Quantitatively, the halo loses \(>90\%\) its initial
mass, as listed in Table~\ref{tbl:scl_sim_results}. However, the inner
structure is not expected to be affected, as the Jacobi radius is over
3kpc, outside of the initial and final \(\rmax\). Thus, tides may remove
significant amounts of mass, but mostly from the outer halo.

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_sim_images.png}}
\caption[Sculptor simulation snapshots]{Images of the dark matter
evolution over a selection of past apocentres and the present day.
Limits range from -150 to 150 kpc in the \(y\)--\(z\) (\(\sim\)orbital)
plane, and the colourscale is logarithmic spanning 5 orders of magnitude
between the maximum and minimum values. The green dot represents the
final expected position of the galaxy and the solid and dotted grey
curves represents the orbit over one previous or future radial
oscillation respectively.}\label{fig:scl_sim_images}
\end{figure}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_tidal_track.png}}
\caption[Sculptor tidal tracks]{Dark matter evolution for the
\smallperi{} model of Sculptor. Blue solid and orange dashed lines show
the initial and final circular velocity profiles. The points represent
the evolution of the maximum circular velocity, and the dotted black
line shows the tidal track from \citet{EN2021}. To calculate the
velocity profiles, unbound particles are iteratively removed,
recalculating the potential at each step assuming spherical symmetry.
\textbf{mark jacobi radius and stellar contributions\ldots, maybe sigmav
/ r\_h}}\label{fig:scl_tidal_track}
\end{figure}

\begin{table*}[t]
\centering
\caption[Simulation results for Sculptor’s dark matter]{The orbital and dark matter properties for the simulation of Sculptor. The random samples column shows the distributions from point orbits, and the \smallperi{} column contains the results from the N-body simulation. }
\label{tbl:scl_sim_results}
\begin{tabular}{lll}
\toprule
Property & random samples & \smallperi{}\\
\midrule
pericentre & $53\pm3$ & 29\\
apocentre & $102\pm3$ & 94.4\\
time of last pericentre / Gyr & $-0.45 \pm 0.2$ & -0.47\\
number of pericentres & 6 & 6\\
Jacobi radius / kpc & $4.5 \pm 0.3$ & 3.5\\
Jacobi radius / arcmin & $186\pm12$ & 101\\
final heliocentric distance / kpc & $83.2\pm2$ & 81.6\\
$\V_\textrm{max, f} / \V_\textrm{max, i}$ &  & 0.695\\
$r_\textrm{max, f} / r_\textrm{max, i}$ &  & 0.406\\
fractional final bound mass &  & 0.0893\\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Evolution of Sculptor's
stars}\label{evolution-of-sculptors-stars}

Tides minimally affect the stellar component of Sculptor in the
\smallperi{} orbit. In Fig.~\ref{fig:scl_smallperi_i_f}, the projected
stellar distribution displays no prominent distortions, and the radial
density profile is nearly unchanged. Only at a surface density
\(\sim10^8\) times fainter than the centre do some faint tidal features
emerge. The total stellar mass lost corresponds to \(\sim 10\) stars in
total (see Table~\ref{tbl:scl_sim_results})---a formidable challenge
with the best of observations.

Sculptor may not have been born as an exponential galaxy, and more
extended stellar profiles would be more sensitive to tides. As an
approximation of the present-day extended density profile, we initialize
a model with Plummer stars instead. As shown in
Fig.~\ref{fig:scl_smallperi_plummer_i_f}, the Plummer model loses more
stellar mass and forms more luminous tidal tails. Observations reaching
surface densities \(\sim10\) times fainter than our data could reveal a
stream in this case. Nevertheless, the radial extent probed by our data
remains nearly unchanged by tidal evolution.

The Jacobi and break radii further support that tidal effects should not
be apparent in the observed stellar component. As calculated for this
model, (see
Tables~\ref{tbl:scl_sim_results}, \ref{tbl:scl_sim_stars_results}), the
break and Jacobi radii both fall outside of \(\sim 100\) arcminutes for
either stellar component. Indeed, the stellar component only begins to
deviate from an exponential profile around this break radius
(Figs.~\ref{fig:scl_smallperi_i_f}, \ref{fig:scl_smallperi_plummer_i_f}).
Given that no orbits of Scl in this potential produce significantly
smaller break or Jacobi radii, it is unlikely any other orbit would
produce a significant, observable density excess.

Table~\ref{tbl:scl_sim_stars_results} quantifies the evolution of
stellar properties. The stellar velocity dispersion decreases by only
\(\sim1\,\kms\) and the half-light radius expands by \(\sim 10\%\). This
is consistent with adiabatic expansion due to the reduction of the total
mass \citep[e.g.,][]{stucker+2023}. In addition, the break and Jacobi
radii are \(\gtrsim 100\) arcminutes on the sky---tidal signatures would
be beyond the reach of our data. Altogether, Galactic tides negligibly
impact Scl's stellar component.

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_smallperi_i_f.pdf}}
\caption[Sculptor initial and final density profiles]{The tidal effects
on Scl's stellar component, for the \smallperi{} orbit with the fiducial
halo and exponential stars with \(R_s=0.10\,\kpc\). \textbf{Top:} the
initial (left) and final (right) 2D projected density of stars on the
sky. The solid circle marks \(6R_h\), the dotted circle the break
radius, and the (dotted) line the past (future) orbit. \textbf{Bottom:}
The initial (dotted) and final (solid) stellar density profiles as
compared to the observed stellar density profile. Arrows mark the break
and Jacobi radii (Eqs.~\ref{eq:r_break}, \ref{eq:r_jacobi})
.}\label{fig:scl_smallperi_i_f}
\end{figure}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_plummer_i_f.pdf}}
\caption[Sculptor Plummer initial and final density profiles]{Similar to
Fig.~\ref{fig:scl_smallperi_i_f} except for Plummer initial stars with
\(R_h = 0.20\,\kpc\). While a faint stream may be visible with deeper
observations, effects on the stellar profile are
minimal.}\label{fig:scl_smallperi_plummer_i_f}
\end{figure}

\begin{table*}[t]
\centering
\caption[Simulation results for Sculptor’s stars]{The present-day stellar properties for the simulations of Sculptor. In each row, we have the initial stellar velocity dispersion (within 1kpc), the final velocity dispersion, the fraction of stellar mass unbound, the initial half-light radius, the final half-light radius, and the break radius in arcmin and kpc (Eq. \ref{eq:r_break}). }
\label{tbl:scl_sim_stars_results}
\begin{tabular}{lll}
\toprule
Property & Exponential & Plummer\\
\midrule
$\sigma_{\V, i}\,/\,\kms$ & 9.8 & 10.7\\
$\sigma_{\V, f} \,/\,\kms$ & 8.8 & 9.4\\
fractional stellar mass loss & $2.1\times 10^{-6}$ & $0.024$\\
$R_{h, i}\,/\,\kpc$ & 0.169 & 0.202\\
$R_{h, f}\,/\,\kpc$ & 0.189 & 0.227\\
break radius / arcmin & $98$ & $105$\\
break radius / kpc & 2.3 & 2.5\\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Orbital effects of the LMC}\label{sec:scl_lmc}

The Milky Way isn't the only galaxy in town. Recently, work has shown
that the infall of the LMC may substantially affect the Milky Way system
\citep[e.g.,][]{erkal+2019, cautun+2019, garavito-camargo+2021, vasiliev2023}.
With a mass up to one fifth of the MW \citep[e.g.,][]{penarrubia+2015},
the LMC infall affects conclusions about the MW properties and the
orbits of satellites \citep[see
e.g.,][]{patel+2020, battaglia+2022, correamagnus+vasiliev2022}. In this
section, we examine how the LMC affects the orbital history of Sculptor.

We use the \texttt{L3M11} model of the MW and LMC potential from
\citet{vasiliev2024}. The \texttt{L3M11} potential is an evolving
multipole approximation of an N-body simulation including a live MW and
LMC dark matter halo. The potential includes a static MW bulge and disk,
evolving MW and LMC halos, and the MW reflex motion. In their
simulation, the MW was initially a NFW halo with \(r_s=16.5\,\)kpc and
\(M_{\rm 200}= 98.4\times10^{10}\Mo\) , and the LMC a NFW halo with
\(r_s=11.7\) and \(M_{200} = 24.6 \times 10^{10} \Mo\). The total
\texttt{L3M11} MW mass is lighter than our initial \citet{EP2020}
potential.

The inclusion of the LMC reshapes Scl's orbital history, as shown in
Fig.~\ref{fig:scl_lmc_orbits_effect}. In the MW-only potential, Scl's
orbit is typical of a long-term MW satellite. However, Scl's passage by
the LMC \(~\sim0.1\,\Gyr\) ago affects the long-term orbit---Scl instead
reaches an apocentre of nearly \(300\,\kpc\). Scl may even be on first
infall, depending on the MW and LMC mass. Scl's orbit is
counter-rotating to the LMC, so Scl is unlikely to be a LMC satellite.

The timing of the LMC encounter implies a break radius consistent with
the beginning of the observed density excess. We select the orbit with
the \(6\sigma\)-smallest LMC pericentre. The orbit is selected following
the procedure in Section~\ref{sec:orbital_estimation} except with
observational uncertainties doubled. Results are similar when selecting
for a small MW pericentre instead. This \texttt{LMC-flyby} orbit, is
integrated back in time \(2\,\Gyr\) ago to isolate recent tidal effects.
We modify Scl's initial halo to have \(\rmax = 2.5\,\kpc\) and
\(\vmax = 25\,\kms\), slightly reducing the initial stellar velocity
dispersion. Fig.~\ref{fig:scl_lmc_orbits_effect} shows this selected
orbit in black and Table~\ref{tbl:orbit_ics} records the initial
conditions.

\begin{table*}[t]
\centering
\caption[Orbits and results for Scl in the LMC potential.]{The orbital properties and dark matter evolution for the models including an LMC. Similar to Table \ref{tbl:scl_sim_results} except quantities with respect to the LMC are in parentheses. }
\label{tbl:scl_lmc_sims}
\begin{tabular}{lll}
\toprule
Property & random samples & \texttt{LMC-flyby}\\
\midrule
pericentre & $44\pm 3$ ($29 \pm 2$) & 39 (20)\\
apocentre & $218 \pm 8$ & –\\
time of last pericentre / Gyr & $-0.38\pm0.01$ (-0.11) & -0.33 (-0.10)\\
number of pericentres & 1 (2) & 1 (1)\\
Jacobi radius / kpc & $4.1\pm0.3$ ($4.5\pm0.2$) & 2.8 (2.6)\\
Jacobi radius / arcmin & $168 \pm 11$ ($186\pm6$) & 132 (121)\\
final heliocentric distance / kpc & $83.2\pm2$ & 72.9\\
$\V_\textrm{max, f} / \V_\textrm{max, i}$ &  & 0.928\\
$r_\textrm{max, f} / r_\textrm{max, i}$ &  & 0.763\\
fractional final bound mass &  & 0.5402\\
\bottomrule
\end{tabular}
\end{table*}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_lmc_xyzr_orbits.png}}
\caption[Sculptor orbits with LMC]{Similar to Fig.~\ref{fig:scl_orbits}
except for orbits with (orange) and without (green lines) the inclusion
of an LMC (blue line) in the potential. The bottom row additionally
shows the distance between Scl and the LMC over
time.}\label{fig:scl_lmc_orbits_effect}
\end{figure}

\subsection{Tidal effects from the
LMC}\label{tidal-effects-from-the-lmc}

The combined tidal effect of the MW and LMC are even weaker for Scl than
in the MW-only case. Fig.~\ref{fig:scl_lmc_sim_images} shows the dark
matter evolution of Scl and the passage of the LMC. With only one MW
pericentre, Scl's dark matter is less disrupted than the previous
MW-only model. The subsequent LMC passage modifies Scl's orbit but has
otherwise little effect. As a result, the dark matter structure evolves
mildly and \(\sim 50\%\) of mass remains bound
(Table~\ref{tbl:scl_lmc_sims}).

Correspondingly, the stellar component is nearly unchanged.
Fig.~\ref{fig:scl_lmc_i_f} shows the projected stellar distributions and
density profiles of this model. While the break radius is within the
observed density profile, tidal effects are too weak to be detectable.
Structural properties of the stars similarly evolve little
(Table~\ref{tbl:scl_lmc_sim_stars}).

While the instantaneous tidal force from the LMC is larger than the MW,
Scl does not experience the LMC tidal field long enough to display
disturbances. Scl and the LMC have a similar orbital plane, but orbit in
opposite directions (Fig.~\ref{fig:scl_lmc_orbits_effect}). As a result,
the interaction between Scl and the LMC is brief as their relative
velocity is \(\sim 400\,\kms\) at closest approach.

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_lmc_sim_images.png}}
\caption[Sculptor simulation snapshots with LMC]{Similar to
Fig.~\ref{fig:scl_sim_images} except for the case where the potential
includes an LMC. The current position and path of the LMC are
represented by the green dot and line respectively. We also plot the
full orbit (over the past 2Gyr) for both Scl and the LMC, as less than
one radial period happens over this time
frame.}\label{fig:scl_lmc_sim_images}
\end{figure}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/scl_lmc_i_f.pdf}}
\caption[Sculptor initial and final density with LMC]{Similar to
Fig.~\ref{fig:scl_smallperi_i_f} except for the \texttt{LMC-flyby}
model. The Jacobi and break radii here are calculated with respect to
the LMC, the corresponding radii with respect to the MW are larger. With
only one MW pericentre and a recent, rapid LMC encounter, tidal forces
do not appear to affect the stellar distribution. \textbf{todo fix break
radius position, orbits as arrows\ldots{}}}\label{fig:scl_lmc_i_f}
\end{figure}

\begin{table*}[t]
\centering
\caption[Simulation results for Sculptor’s stars in the LMC+MW potential]{Similar to Table \ref{tbl:scl_sim_stars_results}, but for the properties of the stellar components of the \texttt{LMC-flyby} model of Sculptor. }
\label{tbl:scl_lmc_sim_stars}
\begin{tabular}{lll}
\toprule
Property & Scl: LMC-exponential & LMC-Plummer\\
\midrule
$\sigma_{\V, i}\,/\,\kms$ & 9.0 & 9.4\\
$\sigma_{\V, f} \,/\,\kms$ & 8.8 & 9.2\\
fractional stellar mass loss & $<10^{-12}$ & 0.0013\\
$R_{h, i}$ / kpc & 0.186 & 0.201\\
$R_{h, f}$ / kpc & 0.189 & 0.205\\
break radius & $77'$, 1.6 kpc & $80'$, 1.7 kpc\\
LMC break radius & $23'$, 0.49 kpc & $24'$, 0.52 kpc\\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Summary}\label{summary}

We find, including only the MW potential, that tides only remove dark
matter from the outskirts of Scl. The central cusp and compact stellar
distribution are resilient to tides and any tidal effects are
well-outside the reach of current observations. We then find that the
LMC strongly perturbs Scl's orbit---in this case, Scl may be on first
infall. However, with only 1 pericentre each for the LMC and MW, the
combined tides are weaker than for our initial model. In either case, we
conclude that tides are unlikely to affect Sculptor's stellar component.

\section{Tidal effects on Ursa Minor}\label{tidal-effects-on-ursa-minor}

\subsection{Evolution of Ursa Minor's dark matter
halo}\label{evolution-of-ursa-minors-dark-matter-halo}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/umi_sim_images.png}}
\caption[Ursa Minor simulation snapshots]{Similar to
Fig.~\ref{fig:scl_sim_images} except for Ursa Minor on the \smallperi{}
orbit. Dark matter evolution is more dramatic than for
Scl.}\label{fig:umi_sim_images}
\end{figure}

\begin{table*}[t]
\centering
\caption[Simulation results for Ursa Minor’s dark matter]{The present-day properties for Ursa Minor’s final dark matter halo. See Table \ref{tbl:scl_sim_results} for details. }
\label{tbl:umi_sim_results}
\begin{tabular}{lll}
\toprule
Property & Random orbits & \smallperi{}\\
\midrule
pericentre & $37\pm3$ & 30\\
apocentre & $83 \pm 4$ & 75\\
time of last pericentre & $-0.97 \pm 0.07$ & -0.80\\
number of pericentres & $\sim 8$ & 8\\
Jacobi radius / kpc & $3.7 \pm 0.3$ & 2.9\\
Jacobi radius / arcmin & $184 \pm 12$ & 156\\
final heliocentric distance & $70.1 \pm 3.6$ & 64.7\\
${\vmax}_f / {\vmax}_i$ &  & 0.511\\
${\rmax}_f / {\rmax}_i$ &  & 0.249\\
fractional dm final mass &  & 0.035\\
\bottomrule
\end{tabular}
\end{table*}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/umi_tidal_track.png}}
\caption[Ursa Minor tidal tracks]{Similar to
Fig.~\ref{fig:scl_tidal_track} except for Ursa Minor. Ursa Minor looses
substantially more mass than Sculptor.}\label{fig:umi_tidal_track}
\end{figure}

The tidal evolution of Ursa Minor is similar to Sculptor in the MW-only
potential. Fig.~\ref{fig:umi_sim_images} shows snapshots of the DM
evolution. UMi loses significantly more DM mass than Scl, forming
substantial dark matter streams encircling the MW several times.

UMi only retains 3\% of its total mass after 9 Gyr
(Table~\ref{tbl:umi_sim_results}). As a result, the final dark matter
component is much smaller than the initial, but still evolves along the
predicted tidal track Fig.~\ref{fig:umi_tidal_track}. Despite the more
substantial tidal evolution, the Jacobi radius is still large, lying at
\(\sim 4\,\kpc\), well beyond the final \(\rmax\).

Because of UMi's mass loss, the orbit deviates substantially from a
point orbit. Through our orbit-adjustment procedure in
Section~\ref{sec:orbit_corrections}, we recover nearly exactly the
present-day position of Ursa Minor by changing the initial positions by
\(20\,\kpc\) and \(\sim 9\,\kms\). These adjustments do not
significantly affect the qualitative structure or pericentre of the
orbit.

\subsection{Evolution of Ursa Minor's
stars}\label{evolution-of-ursa-minors-stars}

\begin{table*}[t]
\centering
\caption[Simulation results for Ursa Minor’s stars]{Similar to Table \ref{tbl:scl_sim_stars_results}, the present-day stellar properties for the simulation of Ursa Minor for exponential and Plummer stars. }
\label{tbl:umi_sim_stars_results}
\begin{tabular}{lll}
\toprule
Property & smallperi-exp & smallperi-Plummer\\
\midrule
$\sigma_{\V, i}\,/\,\kms$ & 10.0 & 10.9\\
$\sigma_{\V, f}\,/\,\kms$ & 8.2 & 8.5\\
fractional stellar mass loss & $0.00015$ & 0.039\\
$R_{h, i}\,/\,\kpc$ & 0.135 & 0.151\\
$R_{h, f}\,/\,\kpc$ & 0.169 & 0.191\\
break radius & 197 arcmin, 3.7 kpc & 204 arcmin, 3.8 kpc\\
\bottomrule
\end{tabular}
\end{table*}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/umi_smallperi_i_f.pdf}}
\caption[Ursa Minor simulated density profiles]{The tidal effects on the
stellar surface density of Ursa Minor for exponential stars on the
\smallperi{} orbit.}\label{fig:umi_smallperi_i_f}
\end{figure}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/umi_plummer_i_f.pdf}}
\caption[Ursa Minor Plummer model density]{The tidal effects on the
stellar surface density of Ursa Minor for Plummer stars on the
\smallperi{} orbit.}\label{fig:umi_plummer_i_f}
\end{figure}

Tidal features in UMi's stellar component are still extremely faint,
becoming apparent only outside 100 arcminutes in
Fig.~\ref{fig:umi_smallperi_i_f}. Tides also only moderately affect the
observed size and velocity dispersion of Ursa Minor
(Table~\ref{tbl:umi_sim_stars_results}). For exponential initial
conditions, tidal effects are unlikely to ever be observable in the near
future.

Similar to Scl, we consider a model where UMi's stars are initially a
more extended Plummer profile, resembling the present-day density
profile. The stellar evolution of the Plummer stellar component is
similar to the exponential stellar profile (\citet{ig:umi_plummer_i_f}).
However, because there are more loosely-bound stars, the Plummer model
looses nearly 4\% of its initial stellar mass to tides
(Table~\ref{tbl:umi_sim_stars_results}), and tidal features may be
detectible if we measure densities 2 orders of magnitude fainter than
our present data. We show the properties of a stream in the Appendix
(Fig.~\ref{fig:umi_tidal_stream}), but such a stream is unlikely to be
observable in the near future.

\subsection{Effects of the LMC}\label{effects-of-the-lmc}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/umi_lmc_xyzr_orbits.png}}
\caption[Ursa Minor orbits with LMC]{Orbits of Ursa Minor with (orange)
and without (green) an LMC. The final positions of Ursa Minor and the
LMC are plotted as scatter points and the solid blue line represents the
LMC trajectory. Note that the LMC mostly increases Ursa Minor's
pericentres and apocentres.}\label{fig:umi_orbits_lmc}
\end{figure}

Fig.~\ref{fig:umi_orbits_lmc} shows the effects of including an LMC on
the orbit of Ursa Minor. Predominantly, the effect is to increase the
orbital period, apocentre and pericentre. However, the orbit remains in
a similar plane and with similar morphology. As UMi is on the opposite
side as the Galaxy of the LMC, this is not unexpected.

The deviation from the MW-only orbit is mostly due to the LMC-induced
reflex motion of the Milky Way. Because the MW centre, and our origin,
is accelerated towards the LMC, UMi experiences a fictitious
acceleration opposite to the acceleration experienced by the MW centre,
causing the orbit to increase in characteristic radius.

\subsection{Summary}\label{summary-1}

While tides affect UMi more strongly than Scl, the tidal effects are
insufficient to reshape the observed stellar density profile. Faint
tidal tails may be observable with deeper data. Finally, including the
LMC in the potential further weakens the tides experienced by UMi.

\section{Modeling uncertainties}\label{modeling-uncertainties}

\subsection{Modelling assumptions}\label{modelling-assumptions}

As Section~\ref{sec:results} shows, tides marginally affected the
stellar components of Scl and UMi, even under extreme orbits. While
Section~\ref{sec:results} only presents select models, alternative
initial conditions do not affect our qualitative conclusions on tidal
effects expected for Scl and UMi in the Galactic (and LMC) potential.

We neglect baryonic physics. Since Scl and UMi have predominantly stars
older than \(\sim 9\) Gyr
\citep{carrera+2002, deboer+2011, weisz+2014, delosreyes+2022, sato+2025},
gas dynamics are unlikely to affect recent evolution. A collisionless
dark-matter only simulation should be an excellent approximation.

Cored or less concentrated dark matter halos disrupt quicker
\citep[e.g.,][]{stucker+2023}. Our fiducial UMi halo in particular is
among the least concentrated halos consistent with UMi's velocity
dispersion. Although Scl's fiducial halo is more concentrated, less
concentrated and cored halos evolve similarly (see
Section~\ref{sec:extra_results}).

Galaxies are rarely perfect isotropic spheres. Sculptor and Ursa Minor
are elliptical, and halos are expected to be radially anisotropic
\citep[e.g.,][]{navarro+2010}. We test example non-spherical and
anisotropic models in Section~\ref{sec:extra_results}, finding that
these assumptions likely do not alter our conclusions.

A system's observed velocity dispersion directly constrains the mean
density within \(R_h\) \citep[e.g.,][]{wolf+2010}. Thus, the tidal force
required to disrupt the stellar component does not strongly depend on
the inner halo structure (via the Jacobi radius). Alternative initial
conditions may influence the total mass evolution but would produce a
similar final stellar structure.

\subsection{Orbital uncertainties}\label{orbital-uncertainties}

The long-term orbits of satellites are uncertain. Analytic Milky Way
potentials neglect many unknown details, including triaxiality, halo
twisting, mass evolution, and substructure. As a result, calculated
orbits may diverge significantly from the true orbits of satellites
\citep[e.g.,][]{dsouza+bell2022}. The mass-growth of the Milky Way and
dynamical friction implies that orbits are typically less bound in the
past (less affected by tides). However, orbital energy and angular
momentum are not conserved in cosmological simulations. Orbits in
analytic potentials may overestimate the pericentre and underestimate
the maximum tidal stress \citep[although typically not by enough to
change our conclusions,][]{santistevan+2023, santistevan+2024}.

Fig.~\ref{fig:scl_orbit_lmc_uncert} illustrates the uncertainties of the
long-term orbits of Scl and UMi. Scl distance could range from
4--\(300\,\kpc\) any time more than 4Gyr ago. And, if the LMC had a
previous pericentre, UMi may have been an LMC satellite
\citep[see][]{vasiliev2024}. Notably, these potentials are still
idealized in many ways.

Motivated by Fig.~\ref{fig:scl_orbit_lmc_uncert}, we examine an extreme
pericentre of \(4\,\kpc\) of Scl with the MW in Appendix
\ref{sec:extra_results}, finding it still insufficient to produce the
observed density profile. We conclude our simulated orbits represent
reasonable extremes for \emph{recent} tidal effects. Past encounters
with the LMC are revisited below as a form of ``pre-processing'' in
Section~\ref{sec:stellar_halos}.

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{/Users/daniel/thesis/figures/scl_lmc_orbits_mass_effect.png}}
\caption[Sculptor Orbits with LMC]{The long-term orbital history of
Sculptor (\textbf{top}) and UMi (\textbf{bottom}) are uncertain. In both
panels, light, transparent lines represent randomly-sampled orbit of the
satellites (alla ref) in three different LMC/MW mass models from
\citet{vasiliev2024}. The LMC orbits are in solid, thick lines of the
corresponding colour. The L2M11 has a lighter LMC mass, and the L3M10
model has a lighter MW mass than our fiducial L3M11 LMC
model.}\label{fig:scl_orbit_lmc_uncert}
\end{figure}

\subsection{Summary}\label{summary-2}

While the long-term tidal evolution is unconstrained, we conclude that
our models are reasonable representations of recent tidal effects. As a
result, recent tides are unlikely to affect the stellar distributions of
Sculptor or Ursa Minor.