### 2025-10-02 thesis/discussion.tex
discussion revisions

In the previous chapters, we have shown that Sculptor and Ursa Minor
have more extended density profiles compared to other MW classical
satellites, and that recent tides are an unlikely explanation for such
extended density profiles. We therefore suggest that Scl and UMi host a
``stellar halo''---an additional, extended population of stars. With our
data, other classical dwarfs lack evidence for similar stellar halos.
The absence of similar halos in other classical dwarfs suggests Scl and
UMi have followed a divergent evolutionary path.

In this chapter, we place our results in context with past work. We
review both work considering tidal effects on Scl and UMi and detections
of peculiarities. We then consider possible origins of stellar halos for
these galaxies. Scenarios include induced star formation forming
multiple populations, past mergers depositing stars into the outskirts,
and even departures from the standard dwarf galaxy structure. We later
discuss how future observations and simulations may distinguish halo
formation scenarios. Finally, we summarize the main conclusions of this
thesis.

\section{Comparison with prior work}\label{comparison-with-prior-work}

\subsection{The role of tides}\label{the-role-of-tides}

Tidal effects have been long invoked to explain features of dwarf
galaxies e.g. \ref{sec:scl_umi_obs_tides},
\citep{tsujimoto+shigeyama2002, mayer+2001a}. Yet, growing work suggests
tides may not be as commonplace as suspected. For example,
\citet{read+2006} suggest that the absence of rising radial velocity
dispersion profiles in dwarfs implies a lack of tidal disruption.
\citet{penarrubia+2009}, introducing the break radius as a diagnostic
(Eq.~\ref{eq:r_break}), showing that most satellite orbits were
inconsistent with observable tidal effects. Recently,
\citet{pace+erkal+li2022} use a Jacobi radius-based criterion
(Eq.~\ref{eq:r_jacobi}), reaching a similar conclusion. With the
exception of the Sagittarius dSph, these results suggest tides have mild
influence on the stellar components of classical dwarfs.

A few works have modelled Scl or UMi specifically, finding similar
conclusions. \citet{iorio+2019} applied idealized N-body simulations to
study tidal effects on Scl. They similarly found weak tidal effects,
even for a dark-matter-free model. Most recently,
\citet{tchiorniy+genina2025} also use idealized simulations tuned to
five classical dwarfs with a focus on the inner density, concluding that
tides do not strongly affect equilibrium assumptions or the stellar
component.

Our work extends these models, adding updated structural properties, a
broader range of orbital histories, and the influence of the LMC.
Despite these considerations, we still reach similar conclusions: tidal
effects do not shape Scl or UMi's stellar component. A non-tidal
explanation likely underlies the origin of these galaxies' density
excess.

\subsection{Peculiarities in the
galaxies}\label{peculiarities-in-the-galaxies}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{/Users/daniel/thesis/figures/scl_umi_fe_h_gradient.pdf}}
\caption[Metallicity gradients in Scl and UMi]{Density profiles and
metallicity gradients in Scl (left) and UMi (right). In the top panel,
we plot the J+24 density profiles as function of elliptical radius in
black, and the two components of a double-exponential density fit as
curved, transparent lines. The vertical lines represent the transition
from the inner to the outer exponential profile. \textbf{Bottom}:
Metallicies as a function of elliptical radius for stars from
\citet{tolstoy+2023}; \citet{apogee}; and \citet{sestito+2023} (left)
and \citet{pace+2020}; \citet{apogee}; and \citet{sestito+2023b}
(right)}\label{fig:metallicity_gradients}
\end{figure}

Besides extended stellar density profiles, Scl and UMi display other
peculiarities which may hold clues to their formation. Scl and UMi both
hosts at least two chemodynamic populations, as revealed through their
photometric or metallicity-velocity structure
\citep{tolstoy+2004, battaglia+2008, pace+2020}. The inner, compact
population is younger, more metal-rich, and dynamically colder, whereas
the outer, extended population older, more metal-poor, and dynamically
hotter.

Could these internal stellar populations map to the observed density
excess? In Scl, both populations have scale radii likely smaller than
the half-light radius. But, the third, more tentatively detected
population in \citet{arroyo-polonio+2024} would reside within the
stellar halo. In Ursa Minor, the transition between the metal-rich and
metal-poor stellar components occurs at \(\sim 30\,\)arcmin
\citep{pace2020}, consummate with the beginning of the density excess.
This metal-poor population may thus be related to the stellar halo of
Ursa Minor. Fig:metallicity\_gradients shows the radial metallicity
distribution in both galaxies. Scl and UMi both show evidence of a
metallicity gradient, but the outskirts have few stars. Further
observations would be necessary to understand if the stellar halo
represents a chemically distinct population.

Ursa Minor has also shown evidence for possible inner substructure such
as stellar or kinematic ``clumps''
\citep[e.g.,][]{olszewski+aaronson1985, demers+1995, kleyna+1998, battinelli+demers1999, bellazzini+2002}.
While one clump was re-detected kinematically in \citet{pace+2014},
\citet{munoz+2018} find no evidence of substructure with modern
photometry. In any case, the survival of a cold clump depends on the
nature of dark matter, disrupting quickly in a \LCDM{} halo
\citep{kleyna+2003, lora+2012}.

Given their multiple stellar populations and possible inner
substructure, this hints that Scl and UMi's distinct structure today
arises from an unusual past. A mechanism forming a stellar halo could
also explain multiple populations.

\section{Forming a stellar halo}\label{sec:stellar_halos}

As we disfavour a tidal origin for Scl and UMi's extended density
profiles, we now consider alternative processes which may form a stellar
halo in these dwarfs.

\textbf{Episodic star formation.} Star formation may quench and
reignite, creating successive stellar generations with differing
distributions. External star formation triggers include tidal
compression \citep{mayer+2001a, dong+lin+murray2003}, collisions with
gaseous filaments \citep{genina+2019}, perturbations from dark halos
\citep{starkenburg+helmi+sales2016}, or shocks with the MW corona
\citep{wright+2019}. More common mechanisms, like feedback or
reionization-driven quenching, may also form multiple stellar
generations
\citep{kawata+2006, benitez-llambay+2015, revaz+jablonka2018}. However,
such processes would not explain why stellar halo formation may be
non-universal. If a stellar halo was formed in an older, induced burst
of star formation, then the star formation history should contain
evidence of a corresponding burst.

\textbf{Major mergers.} Dwarf galaxy mergers may be relatively common
\citep{deason+wetzel+garrison-kimmel2014}. After a merger, stars from
the lower-mass galaxy are preferentially dispersed, forming a stellar
halo and population gradient \citep{benitez-llambay+2016, deason+2022}.
If the galaxies contains gas, the merger can trigger new star formation,
forming a younger population of stars \citep[e.g.,][]{genina+2019}. A
few local dwarfs are suspected to have undergone a major merger,
including Tucana II, Andromeda II, and Phoenix
\citep{lokas+2014, fouquet+2017, tarumi+2021, cardona-barrero+2021, querci+2025}.
A galaxy having undergone a major merger should show evidence of at
least two or three populations from distinct origins.

\textbf{Minor mergers / accretion}. Just as the Milky Way's halo is
believed to be built from many minor mergers, dwarf galaxies may build
stellar halos composed of even fainter satellites
\citep{ricotti+polisensky+cleland2022}. Few examples are known, but a
stream around And II may be an instance of dwarf accretion
\citep{amorisco+evans+vandeven2014, roederer+2023}. In this scenario,
the stellar halo's would contain chemical signatures from several
distinct ultra-faint dwarf galaxies.

\textbf{Tidal preprocessing}. Some dwarf galaxies may have been tidally
``preprocessed'' by a massive satellite like the LMC
\citep[e.g.,][]{santistevan+2023, riley+2024}. Tidal preprocessing
redistributes already-present stellar populations without creating new
populations, and may mimic a stellar halo. Key evidence suggesting
preprocessing may include distant dwarf stars or stellar streams.

\textbf{Dark-matter free dwarfs.} Tidal dwarfs form in gas-rich tidal
streams created during the merger of two galaxies
\citep[e.g.,][]{mirabel+dottori+lutz1992, bournaud+duc2006}. Without a
dark matter halo, tidal dwarfs may be more susceptible to tides, forming
extended density profiles and inflating the velocity dispersion
\citep{casas+2012, yang+2014, wang+2024a}.

Similarly, in Modified Newtonian Dynamics (MOND), galaxies do not
contain dark matter. Instead the gravitational force law is altered to
explain rotation curves. Dwarfs would similarly experience stronger
tidal effects in MOND \citep{mcgaugh+wolf2010, brada+milgrom2000}.

In both cases, tides more plausibly produce extended density profiles.
However, recovering the observed velocity dispersions often requires
ongoing tidal disruption
\citetext{\citealp{mcgaugh+wolf2010}; \citealp[but see
also][]{sanchez-salcedo+hernandez2007}}. Current data do not show
evidence of such features (see Section~\ref{sec:extra_rv_models}), so
future data would need to uncover signs of disruption to support this
hypothesis.

\textbf{Dynamical heating.} Old stars in dwarf galaxies may be hotter
than younger stars due to processes including stellar feedback
\citep{stinson+2009, maxwell+2012, el-badry+2016, mercado+2021},
sub-subhalo interactions \citep{penarrubia+2025}, or even fuzzy dark
matter interference fringes
\citep[e.g.,][]{el-zant+2020, duttachowdhury+2023}. However, most of
these processes should operate similarly across dwarf galaxies---stellar
halos should be more common if these processes form substantial halos.

\section{Disentangling the origin of a stellar
halo}\label{disentangling-the-origin-of-a-stellar-halo}

We just reviewed a number of different, possibly-concurrent explanations
for the formation of a stellar halo. While we leave the nature of Scl
and UMi's extended stellar densities an open question, we can discuss
possible clues to different formation scenarios.

\emph{Precise chemistry}. Chemistry, particularly comparing the inner
and outer regions, could test if the halo appears to originate from a
distinct system than the dwarf. This would test if the halo is a
separate chemical population, may reveal the properties of any merged
galaxies, or show if the halo is consistent with originating from the
same galaxy.

\emph{Detailed kinematics}. For models relying on recent tidal
disruption, kinematic disequilibrium features should be visible. These
would appear as velocity gradients, increasing velocity dispersions,
outward-biased moving stars, or non-phase mixed structures
\citep[e.g,][]{kroupa1997, read+2006, sanchez-salcedo+hernandez2007}. In
addition, measurements of the ellipticity and anisotropy would help
verify or alter our understanding of the mass structure and if DM-free
models may be permissible.

\emph{Deep photometry} may find or rule out signs of dynamical
disequilibrium and tidal tails for tidally susceptible models
\emph{(e.g., pre-processing and DM-free dwarfs.)} In addition,
photometry will help constrain the prevalence and nature of extended
features in dwarf galaxies.

\emph{Star formation histories} can be derived through photometry or
chemistry. Evidence of significant star formation episodes or lack
thereof may differentiate scenarios which rely on strong star formation
bursts (e.g.~\emph{episodic star formation history, induced star
formation, and gas-rich mergers.})

Ongoing and future facilities will be essential for testing dwarf galaxy
halo formation. Given the typical faint magnitudes of dSph stars,
chemistry is best accessible through multi-object spectroscopic
instruments on large telescopes (e.g., 4MOST and extremely large
telescopes \citet{skuladottir+2023}). 3D internal kinematics of dwarfs
may require a successor to \emph{Gaia}. However, using JWST and HST
enables precise proper motions for small regions within dwarf galaxies
\citep[e.g.,][]{vitral+2025}, which may be sufficient to constrain
vastly different internal dynamic structures and ongoing tidal
disruption. Finally, ongoing photometric surveys by Rubin Observatory
\citep{ivezic+2019} and the Euclid mission
\citep{euclidcollaboration+2025} will probe even fainter features around
dwarf galaxies, possibly finding more halos, tidal streams, and
improving our understanding of their stellar populations. Altogether,
new surveys will likely uncover novel aspects about the inner workings
and outer halos of dwarf galaxies.

The next generation of dwarf simulations are beginning to answer
questions and unravel processes in the formation of dwarf galaxies.
Notable recent simulation projects include EDGE \citep{agertz+2020} and
LYRA \citep{gutcke+2021}, both of which can simulate the cosmological
formation of isolated dwarf galaxies while resolving individual
(massive) stars. By including more realism (e.g., black holes, magnetic
fields), refining models of poorly understood processes (feedback and
star formation), and improving numerical resolution, simulations will
continue to explore these processes and refine our understanding of
dwarf galaxy formation.

\section{Conclusion and outlook}\label{conclusion-and-outlook}

In this thesis, we have investigated the extended, outer density excess
of Sculptor and Ursa Minor and its possible tidal origin. We show that
the density profiles are robust, implying that Scl and UMi contain a
true density excess relative to other classical dwarfs.

We then investigated if tides are a possible explanation. By modelling
each galaxy based on cosmological initial conditions, we showed that
tides do not strongly affect either galaxy. The LMC changes the orbital
history of Scl and UMi, but tides become even weaker in a combined LMC
and MW potential. We conclude that recent tides are unlikely to shape
the outer radial distribution of stars in Scl and UMi.

Finally, we discuss the implications of a non-tidal density excess in
Scl and UMi. We review mechanisms ranging from episodic star formation
histories, to accretion, to departures from \LCDM{}. While a more
precise explanation awaits, upcoming and ongoing surveys will uncover
the stellar populations within dwarf galaxies in unprecedented detail.
Future work will continue to uncover the cosmic origins of the smallest
structures forming building blocks of galaxies like ours.