### 2025-10-06 thesis/discussion.tex

In the previous chapters, we have shown that Sculptor and Ursa Minor
have more extended density profiles compared to other MW classical
satellites, and that recent tides are an unlikely explanation for such
extended density profiles. We therefore suggest that Scl and UMi may
host an additional, extended population of stars. With our data, other
classical dwarfs lack evidence for similar extended populations. The
absence of similar extended populations in other classical dwarfs
suggests Scl and UMi have followed a divergent evolutionary path.

In this chapter, we place our results in context with past work. We
review both work considering tidal effects on Scl and UMi and detections
of peculiarities. We then consider possible origins of extended stellar
populations for these galaxies. Scenarios include multi-epoch star
formation, past mergers depositing stars into the outskirts, and even
departures from the typical exponential dwarf galaxy structure. We later
discuss how future observations and simulations may distinguish halo
formation scenarios. Finally, we summarize the main conclusions of this
thesis.

\section{Comparison with prior work}\label{comparison-with-prior-work}

\subsection{The role of tides}\label{the-role-of-tides}

Tidal effects have been long invoked to explain features of dwarf
galaxies e.g.~Section \ref{sec:scl_umi_obs_tides},
\citep{mayer+2001a, tsujimoto+shigeyama2002} \textbf{Check this
citation}. Yet, growing work suggests tides may not be as commonplace as
suspected. For example, \citet{read+2006} suggest that the absence of
rising radial velocity dispersion profiles in dwarfs implies a lack of
tidal disruption. \citet{penarrubia+2009}, introducing the break radius
as a diagnostic (Eq.~\ref{eq:r_break}), showed that most satellite
orbits were inconsistent with observable tidal effects. Recently,
\citet{pace+erkal+li2022} use a Jacobi radius-based criterion
(Eq.~\ref{eq:r_jacobi}) to reach a similar conclusion. With the
exception of the Sagittarius dSph, these results suggest tides have mild
influence on the stellar components of classical dwarfs.

A few works have modelled Scl or UMi specifically, finding similar
conclusions. \citet{iorio+2019} applied idealized N-body simulations to
study tidal effects on Scl. They similarly found weak tidal effects,
even for a dark-matter-free model. Most recently,
\citet{tchiorniy+genina2025} also use idealized simulations tuned to
five classical dwarfs with a focus on the inner density, concluding that
tides do not strongly affect equilibrium assumptions or the stellar
component.

Our work extends these models, adding updated structural properties, a
broader range of orbital histories, and the influence of the LMC.
Despite these considerations, we still reach similar conclusions: tidal
effects do not shape Scl or UMi's stellar component. A non-tidal
explanation likely underlies the origin of these galaxies' outer density
excess.

\subsection{Peculiarities in the
galaxies}\label{peculiarities-in-the-galaxies}

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{/Users/daniel/thesis/figures/scl_umi_fe_h_gradient.pdf}}
\caption[Metallicity gradients in Scl and UMi]{Density profiles and
metallicity gradients in Scl (left) and UMi (right). In the top panel,
we plot the J+24 density profiles as function of elliptical radius in
black, and the two components of a double-exponential density fit as
curved, transparent lines. The vertical lines represent the transition
from the inner to the outer exponential profile. \textbf{Bottom}:
Metallicities as a function of elliptical radius for stars from
\citet{tolstoy+2023}; \citet{abdurrouf+2022}; and \citet{sestito+2023a}
(left) and \citet{pace+2020}; \citet{abdurrouf+2022}; and
\citet{sestito+2023b} (right), crossmatched against J+24 as described in
Appendix \ref{sec:extra_rv_models}.}\label{fig:metallicity_gradients}
\end{figure}

Besides extended stellar density profiles, Scl and UMi display other
peculiarities which may hold clues to their formation. Scl and UMi both
hosts at least two chemodynamic populations, as revealed through their
photometric or metallicity-velocity structure
\citep{tolstoy+2004, battaglia+2008, pace+2020}. The inner, compact
population is younger, more metal-rich, and dynamically colder, whereas
the outer, extended population older, more metal-poor, and dynamically
hotter.

Could these internal stellar populations map to the observed density
excess? In Scl, both populations have scale radii likely smaller than
the half-light radius. But, the third, more tentatively detected
population in \citet{arroyo-polonio+2024} would reside within the outer
density excess. In Ursa Minor, the transition between the metal-rich and
metal-poor stellar components occurs at \(\sim 30\,\)arcmin
\citep{pace+2020}, which coincides with the density excess. This
metal-poor population may thus be related to the extended stellar
density of Ursa Minor.

Fig.~\ref{fig:metallicity_gradients} shows the radial metallicity
distribution in both galaxies. To aid in comparison, we fit a
2-component exponential to the density profiles. We also mark where the
densities of each component become equal, the ``transition'' radius. Scl
and UMi both show evidence of a metallicity gradient. But, the relation
between the metallicity gradient and the transition from the inner to
the outer density profile is unclear. UMi also has few measurements past
the transition radius. Further observations would be necessary to
understand if the extended stellar component represents a chemically
distinct population.

Ursa Minor has also shown evidence for possible inner substructure such
as stellar or kinematic ``clumps''
\citep[e.g.,][]{olszewski+aaronson1985, demers+1995, kleyna+1998, battinelli+demers1999, bellazzini+2002}.
While one clump was re-detected kinematically in \citet{pace+2014},
\citet{munoz+2018} find no evidence of substructure with modern
photometry. In any case, the survival of a cold clump depends on the
inner structure of the halo, disrupting quickly in halos with
substructure or inner density cusps but survining longer in smooth,
cored halos \citep{kleyna+2003, lora+2012}.

Given their multiple stellar populations and possible inner
substructure, this hints that Scl and UMi's distinct structure today
arises from an unusual past. For example, a merger may naturally explain
both extended and multiple stellar populations.

\section{Forming an extended stellar
population}\label{sec:stellar_halos}

As we disfavour a tidal origin for Scl and UMi's extended density
profiles, we now consider six alternative processes which may form an
extended stellar population in these dwarfs.

\textbf{Episodic star formation.} Star formation may quench and
reignite, creating successive stellar generations with differing
distributions. External star formation triggers include tidal
compression \citep{mayer+2001a, dong+lin+murray2003}, collisions with
gaseous filaments \citep{genina+2019}, perturbations from dark halos
\citep{starkenburg+helmi+sales2016}, or shocks with the MW corona
\citep{wright+2019}. More common mechanisms, like feedback or
reionization-driven quenching, may also form multiple stellar
generations
\citep{kawata+2006, benitez-llambay+2015, revaz+jablonka2018}. However,
such processes would not explain why extended stellar populations appear
to be non-universal. If an extended population was formed in an older,
induced burst of star formation, then the star formation history should
contain evidence of a corresponding burst.

\textbf{Major mergers.} Dwarf galaxy mergers may be relatively common
\citep{deason+wetzel+garrison-kimmel2014}. After a merger, stars from
the lower-mass galaxy are preferentially dispersed, forming an extended
stellar component and population gradient
\citep{benitez-llambay+2016, deason+2022}. If the galaxies contains gas,
the merger can trigger new star formation, forming a younger population
of stars \citep[e.g.,][]{genina+2019}. A few local dwarfs are suspected
to have undergone a major merger, including Tucana II, Andromeda II, and
Phoenix
\citep{lokas+2014, fouquet+2017, tarumi+yoshida+frebel2021, cardona-barrero+2021, querci+2025}.
A galaxy having undergone a major merger should show evidence of at
least two or three populations from distinct origins.

Just as the Milky Way's halo is believed to be built from many minor
mergers, dwarf galaxies may build accreted ``stellar halos'' composed of
even fainter satellites \citep{ricotti+polisensky+cleland2022}. Few
examples are known, but a stream around And II may be an instance of
dwarf accretion \citep{amorisco+evans+vandeven2014, roederer+2023}. In
this scenario, the extended stellar populations would contain chemical
signatures from several distinct ultra-faint dwarf galaxies.

\textbf{Tidal preprocessing}. Some dwarf galaxies may have been tidally
``preprocessed'' by a massive satellite like the LMC
\citep[e.g.,][]{santistevan+2023, riley+2024}. Tidal preprocessing
redistributes already-present stellar populations without creating new
populations, and may mimic a stellar halo. Key evidence suggesting
preprocessing may include distant dwarf stars or stellar streams.

\textbf{Dark-matter free dwarfs.} Tidal dwarfs form in gas-rich tidal
streams created during the merger of two galaxies
\citep[e.g.,][]{mirabel+dottori+lutz1992, bournaud+duc2006}. Without a
dark matter halo, tidal dwarfs may be more susceptible to tides, forming
extended density profiles and inflating the velocity dispersion
\citep{casas+2012, yang+2014, wang+2024a}.

Similarly, in Modified Newtonian Dynamics (MOND), galaxies do not
contain dark matter. Instead the gravitational force law is altered to
explain rotation curves. Dwarfs would similarly experience stronger
tidal effects in MOND \citep{mcgaugh+wolf2010, brada+milgrom2000}.

In both cases, tides more plausibly produce extended density profiles.
However, recovering the observed velocity dispersions often requires
ongoing tidal disruption
\citetext{\citealp{mcgaugh+wolf2010}; \citealp[but see
also][]{sanchez-salcedo+hernandez2007}}. Current data do not show
evidence of such features (see Appendix \ref{sec:extra_rv_models}), so
future data would need to uncover signs of disruption to support this
hypothesis.

\textbf{Dynamical heating.} Old stars in dwarf galaxies may be hotter
than younger stars due to processes including stellar feedback
\citep{stinson+2009, maxwell+2012, el-badry+2016, mercado+2021},
sub-subhalo interactions \citep{penarrubia+2025}, or even fuzzy dark
matter interference fringes
\citep[e.g.,][]{el-zant+2020, duttachowdhury+2023}. However, most of
these processes should operate similarly across dwarf
galaxies---extended stellar populations should be more common if these
processes form substantial halos.

\section{Disentangling the origin of extended stellar
populations}\label{disentangling-the-origin-of-extended-stellar-populations}

We reviewed a number of different, possibly-concurrent explanations for
the formation of extended populations. While we leave the nature of Scl
and UMi's extended stellar densities an open question, we can discuss
possible clues to different formation scenarios.

\emph{Precise chemistry}. Chemistry, particularly comparing the inner
and outer regions, could test if the halo appears to originate from a
distinct system than the dwarf. This would test if the halo is a
separate chemical population, may reveal the properties of any merged
galaxies, or show if the halo is consistent with originating from the
same galaxy.

\emph{Detailed kinematics}. For models relying on recent tidal
disruption, kinematic disequilibrium features should be visible. These
would appear as velocity gradients, increasing velocity dispersions,
outward-biased moving stars, or non-phase mixed structures
\citep[e.g,][]{kroupa1997, read+2006, sanchez-salcedo+hernandez2007}. In
addition, measurements of the ellipticity and anisotropy would help
verify or alter our understanding of the mass structure and if DM-free
models may be permissible.

\emph{Deep photometry} may find or rule out signs of dynamical
disequilibrium and tidal tails for tidally susceptible models
\emph{(e.g., pre-processing and DM-free dwarfs.)} In addition,
photometry will help constrain the prevalence and nature of extended
features in dwarf galaxies.

\emph{Star formation histories} can be derived through photometry or
chemistry. Evidence of significant star formation episodes or lack
thereof may differentiate scenarios which rely on strong star formation
bursts (e.g.~\emph{episodic star formation history, induced star
formation, and gas-rich mergers.})

Ongoing and future observational surveys will be instrumental for
testing dwarf galaxy halo formation. Because dSph stars are usually
faint, detailed chemical analysis will rely on medium--high resolution
multi-object spectroscopy on large telescopes, such as 4MOST
\citep{skuladottir+2023} or extremely large telescopes. Mapping the 3D
kinematics across a dwarf may require a \emph{Gaia} successor, but
ongoing surveys with JWST and HST can derive precise proper motions
within select regions of dwarf galaxies, able to reconstruct internal
structure and anisotropies \citep[e.g.,][]{vitral+2025}. Current
wide-field photometric surveys by Rubin Observatory \citep{ivezic+2019}
and the Euclid mission \citep{euclidcollaboration+2025} will trace even
fainter features around dwarf galaxies, possibly finding halos, tidal
streams, and unusual stellar populations. Together, these observational
programs promise a new window into the detailed structure, evolution,
and outskirts of dwarf galaxies.

The next generation of dwarf simulations will likewise continue to
refine our understanding of dwarf galaxy formation. Notable recent
simulation projects include EDGE \citep{agertz+2020} and LYRA
\citep{gutcke+2021}, both of which can simulate the cosmological
formation of isolated dwarf galaxies while resolving individual
(massive) stars. Next steps for simulations include improved realism
(e.g., magnetic fields), refining physical models (including feedback
and star formation), and improving numerical resolution. With these
continuing advances, future simulations may shed light on viability of
halo formation scenarios and the details of dwarf galaxy assembly.

\section{Conclusion and outlook}\label{conclusion-and-outlook}

In this thesis, we have investigated the extended, outer density excess
of Sculptor and Ursa Minor and its possible tidal origin. We show that
the density profiles are robust, implying that Scl and UMi contain a
true density excess relative to other classical dwarfs.

We then investigated if tides are a possible explanation. By modelling
each galaxy based on cosmological initial conditions, we showed that
tides do not strongly affect either galaxy. The LMC changes the orbital
history of Scl and UMi, but tides become even weaker in a combined LMC
and MW potential. We conclude that recent tides are unlikely to shape
the outer radial distribution of stars in Scl and UMi.

Finally, we discuss the implications of a non-tidal density excess in
Scl and UMi. We review mechanisms ranging from episodic star formation
histories, to accretion, to departures from \LCDM{}. Future simulations
and observational programs will be able to investigate the nature of
dwarfs with ever increasing detail. Abundant opportunities await to
uncover the cosmic origins and history of dwarf galaxies, the building
blocks of galaxies like ours.