### 2025-10-08 thesis/acknowledgments.tex

While written by my hand, I cannot pretend that I alone created this
work. I am ever thankful and greatly indebted to the multitude of people
who have supported, inspired, and nurtured me as a human being.

To my supervisor, Julio Navarro, for being the critical eye and guiding
reason to this work. I will always approach research differently now.

To Raphaël, who has helped inspire and direct this project. To Isabel
and Ryan.

To Jax, who has been a support along the way, guiding my research
methods and project. Akshara and Ewould.

To all of my new friends I have made along the way. Ruxin, Aviv, Anya,
Pranav, Ted, Adrianne, Will, Maeve, and everyone else I have met. We
have build a wonderful community of students in this department. And to
my friends across UVic, Wyatt, Rosemary, Maggie,

To my past advisors, James who has continued to support me long after
our paths diverged, David and Jennifer for guiding my first research
project.

To Wayne for showing us the joy of astronomy and outreach. To my mentor,
Ebtihal.

To my past roommates, Matthew, Josh, and Bryn, for putting up with my
chaos and our shared time together.

To Fierce Brosnen and our trivia team. Melody, Lily, Gillian, Stephen,
Jillian, and so many more.

To my long-time friends. Nikola and Eric. Mikey and Anthony. Anya, Kaia,
Alyssa, Aahliyah, Cameron. May we continue to find each other even as we
drift apart.

To my musical friends, Adrienne, Alex for our shared joy in music
making. And Steve, for showing me the joy in a simple life and beautiful
music. To my fellow knitters. Anne, Anya, Leise, Wes. To the
international community.

To my parents for their continuous support throghout this project. To my
grandparents for always being there. To my cousins and my extended
family. And to my siblings for always being a source of joy.

To the stars and the moon. To the land of Victoria, and its historical
caretakers. To the waters and beautiful trees, the ocean waves, rocks
and hills and jagged coastlines. To the rich, enveloping forests and
rust-red arbutus tress, scraggly Garry oaks and Camus flowers. I could
not imagine a more beautiful place to study and to always remind my
place in the world.