### 2025-10-08 thesis/extra_numerical_convergence.tex

\chapter{Numerical convergence and
parameters}\label{sec:extra_convergence}

Here, we describe the numerical specifics of our simulation and
analysis. We present convergence tests supporting our choices of
softening and additional parameters. We find that alternate choices of
numerical parameters neither improve convergence nor the resulting
evolution.

Due to the finite resolution of N-body simulations, collisional (close)
encounters between particles are inevitable. Such collisional encounters
are both unphysical (with possible arbitrary acceleration, violating the
``collisionless'' assumption) and computationally expensive (requiring
many small time-steps). As a remedy, many N-body codes use a
``softened'' gravitational force law, where the force weakens when
closer to a particle than a ``softening length''. The choice of
softening length ideally balances resolution and computational speed.

Empirically, \citet{power+2003} suggest that the ideal softening is
\begin{equation}\protect\phantomsection\label{eq:power_softening}{
h_{\rm grav} = 4 \frac{R_{200}}{\sqrt{N_{200}}},
}\end{equation}

where \(h_{\rm grav}\) is the softening length, and \(N_{200}\) is the
number of particles within \(R_{200}\). This choice balances integration
time and only compromises resolution in the collisional regime. For our
isolation halo (\(\rmax=6\,\kpc\), \(\vmax=31\,\kms\)) with \(10^7\)
particles, this works out to be \(0.044\,{\rm kpc}\). Next, we consider
the effect of softening as applied to our isolation halo.

The top panel of Fig.~\ref{fig:methods_convergence} illustrates the
influence of softening length for simulations of Scl's \smallperi{}
orbit with \(10^5\) particles. We show a model with the fiducial
softening length (\(h=0.14\,\kpc\)), and larger and smaller softenings
by a factor \(\sqrt{10}\). The larger softening length is consistent
with Eq.~\ref{eq:power_softening}'s prediction. With a larger softening
length, the halo diverges from the expectation (using \(100\times\) more
particles) across most radii. On the other hand, the fiducial and
smaller softening lengths predict similar final results. As computation
time increases with decreasing softening length, the fiducial softening
length balances efficiency and accuracy for this simulation.

The two other key numerical parameters in \gadget{} are the tree-force
and time-step accuracy parameters. Specifically, \gadget{} opens a node
if \(M\,l/r^3 < \alpha |a|\), where the node's mass is \(M\), the
distance from the particle is \(r\), the side-length is \(l\), and the
particle's total acceleration is \(a\). We adopt \(\alpha =0.005\). We
also elect to use adaptive time stepping with integration accuracy set
to \(\eta=0.01\). (particles must take time-steps smaller than
\(dt < \sqrt{2\,\eta\,h_{\rm grav} / a}\) for acceleration \(a\)). The
lower panel of Fig.~\ref{fig:methods_convergence} tests changes to these
parameters. We include a model with smaller integration accuracy
(\(\eta=0.003\) in ``small timestep'') and a stricter tree-force
tolerance (\(\alpha=0.001\), ``high acc. force''). More precise
tolerances on these parameters do not affect the evolution within the
uncertainties of the final profile.

As demonstrated in this subsection, stricter numerical accuracy and
modified softening lengths do not affect our results. We thus conclude
that our simulations are numerically well-converged (up to the
``convergence'' radius).

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{figures/orbit_converg_methods.png}}
\caption[Isolation method convergence]{A comparison of the final
profiles using different simulation methods for Sculptor's \smallperi{}
model. The benchmark model is our fiducial \(10^7\) particle run, and
all other models use \(10^5\) particles, with their ``converged radius''
marked by the black arrow and softening by the vertical bar(s).
\textbf{Top:} Models with \(\sqrt{10}\) larger and smaller softening
lengths than the fiducial. \textbf{Bottom:} Models with more precise
timestep accuracy and gravitational force accuracy. \textbf{Summary:}
Except for the model with a larger softening length, all simulations
agree within uncertainties.}\label{fig:methods_convergence}
\end{figure}