### 2025-10-08 thesis/main.tex

 %! TeX program = xelatex

\DocumentMetadata{lang=en-CA}

\documentclass[12pt,oneside,letterpaper]{report}

\input{formatting}
\input{acronyms}
\usepackage[single]{accents}

% useful commands
% below are for statement
\newcommand{\schwa}{ə}
\newcommand{\eng}{\ng{}}
%\newcommand{\overcomma}[1]{${\rm \accentset{\mbox{\normalfont ,}}{#1}}$}
\newcommand{\Lekwnen}{L\schwa{}\overcomma{k}\textsuperscript{w}\schwa{}\ng{}\schwa{}n}
\newcommand{\skipline}{\vspace{\baselineskip}}

\def\overcomma#1{%
  \setbox0=\hbox{#1}% Save the base letter in box0.
  \dimen0=\wd0       % Get its width.
  \vbox{%
    \hbox to \dimen0{\hfil,\hfil}% Create an hbox of the same width with a centered comma.
    \nointerlineskip            % Prevent extra vertical space.
    \box0                       % Place the base letter.
  }%
}

% for pandoc
\newcommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\newcommand{\about}[1]{${\sim} #1$}

\title{
    The extended light profiles of the Sculptor and Ursa Minor dwarf galaxies: Nature or tidal nurture?
}
\author{Daniel A. Boyea}

\date{\today}

\begin{document}
% \linespread{1.25}

\pagenumbering{roman}

\makeatletter
\begin{titlepage}
   \begin{center}
       \textbf{\large \@title}\\
        \skipline

        by\\
        \skipline

       \@author\\
       B.Sc. The Ohio State University, 2023\\
       \vspace*{3\baselineskip}
    A Thesis Submitted in Partial Fulfillment of the Requirements for the Degree of\\
    \skipline

    MASTER OF SCIENCE \\
    \skipline

    in the Department of Physics and Astronomy\\
       \vfill
       {\small
       ©\@author, 2025\\
       University of Victoria\\
   }
   \skipline
       {\small
       All rights reserved. This thesis may not be reproduced in whole or in part,
   by photocopy or other means, without the permission of the author.} \\
\skipline
   \end{center}
We acknowledge and respect the \Lekwnen{} (Songhees and X\textsuperscript{w}seps\schwa{}m/
Esquimalt) Peoples on whose territory the university stands, and the
\Lekwnen{} and ${\underaccent{\bar}{\rm W}}$S\'ANE\'C Peoples whose historical relationships with the
land continue to this day. 
\end{titlepage}

\addtocounter{page}{1}

\addcontentsline{toc}{chapter}{Supervisory Committee}
\begin{centering}
\textbf{\@title}\\
\skipline
by\\
\skipline
\@author\\
B.Sc. The Ohio State University, 2023\\
\vspace*{3\baselineskip}
\end{centering}

\subsection*{Supervisory Committee}
\skipline

Dr. Julio Navarro, Supervisor\\
Department of Astronomy \\
\skipline

\noindent Dr. Kim Venn\\
Department of Astronomy\\

\skipline

\noindent External Examiner \\
Outside Department

% Abstract of the paper
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
% context
\input{abstract.tex}

\tableofcontents
\listoffigures
\listoftables

\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
\input{acknowledgments}

%% Lists
%\chapter*{Symbols}
%
%
%\setlength{\tabcolsep}{0pt}
%\begin{longtable}{p{0.2\textwidth} p{0.8\textwidth}}
%$\rho$ & 3-dimensional mass density. \\
%
%$\Sigma$ & 2-dimensional density \\
%\end{longtable}
%
%\setlength{\tabcolsep}{6pt}
%\renewcommand*{\arraystretch}{1}
%
%
%\printglossary[type=\acronymtype,nonumberlist]
%\printglossary
%
%
\newpage

\thispagestyle{empty}
\renewcommand{\epigraphflush}{center}
\vfill
\setlength{\epigraphwidth}{0.5\textwidth}
\epigraph{The sun had fallen beneath the horizon in the east, and the long day was about to end. But tomorrow, the sun would rise in the west again, bathing the world in a kinder light.}{Baoshu, translation by Ken Liu}

\epigraph{
In the land of my birth, there is a hill, and upon that hill is a rock where I have sat and watched a changing world. This rock will not change, I would say. I'll pass from this Earth, but the rock will remain ever watchful of the land.

But my feet kicked the dirt beneath me. Were the granules of soil not once like this rock? Is the rock not a collection of their multitudes?

[We are] neither destroyed nor created, but a collection of dust meant to observe a changing world and then return to dust. 
}{Joe Fisher, {\it Midnight Burger}}
\vfill

\newpage
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\chapter{Introduction}\label{sec:introduction}
\input{introduction}

\chapter{The Light Profiles of the Classical Dwarf Spheroidals}\label{sec:observations}
\input{data}

\chapter{Simulation Methods}\label{sec:methods}
\input{methods}

\chapter{Galactic Tidal Effects on Sculptor and Ursa Minor}\label{sec:results}
\input{results}

\chapter{Discussion \& Conclusions}\label{sec:discussion}
\input{discussion}

\chapter*{Research Acknowledgments}
\addcontentsline{toc}{chapter}{Research Acknowledgments}
\input{research_acknowledgments}

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

\newpage
\bibliographystyle{aasjournal}
\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{main}

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\input{extra_density}
\input{extra_rv_models}
\input{extra_numerical_convergence}
\input{extra_results}

\end{document}