### 2025-10-08 thesis/research_acknowledgments.tex

This work used the DiRAC@Durham facility managed by the Institute for
Computational Cosmology on behalf of the STFC DiRAC HPC Facility
(www.dirac.ac.uk). The equipment was funded by BEIS capital funding via
STFC capital grants ST/K00042X/1, ST/P002293/1, ST/R002371/1 and
ST/S002502/1, Durham University and STFC operations grant ST/R000832/1.
DiRAC is part of the National e-Infrastructure.

Computations were performed on the Niagara supercomputer
\citep{ponce+2019} at the SciNet HPC Consortium \citep{loken+2010}.
SciNet is funded by Innovation, Science and Economic Development Canada;
the Digital Research Alliance of Canada; the Ontario Research Fund:
Research Excellence; and the University of Toronto.

This work has made use of data from the European Space Agency (ESA)
mission \emph{Gaia} https://www.cosmos.esa.int/gaia), processed by the
\emph{Gaia} Data Processing and Analysis Consortium (DPAC),
https://www.cosmos.esa.int/web/gaia/dpac/consortium. Funding for the
DPAC has been provided by national institutions, in particular the
institutions participating in the \emph{Gaia} Multilateral Agreement.

This work has made use of software including Jupyter, astropy, Makie,
Turing, and Pluto.