### 2025-10-17 thesis/acknowledgments.tex

While written by my hand, I cannot pretend that I alone created this
work. I am ever thankful and greatly indebted to the multitude of people
who have supported, inspired, and nurtured me as a human being.

To my supervisor, Julio Navarro, for being the critical eye and guiding
reason to this work. I will always approach research differently now.

To Raphaël, who has helped inspire and direct this project. To everyone
who has helped provide me opportunities to share work, and share
wonderful ideas with me. Isabel, Jens, Eduardo, Jorge, David, Justin,
Eugene, Azi, José, and many others.

To Jax, who has been a support along the way, guiding my research
methods and project.

To all of my new friends I have made along the way. Ruxin, Aviv. To
Pranav, Ted, Jono, Adrianne, Praneet, Will, Maeve, Ewoud. To Akshara for
all your personal and professional advice. We have build a wonderful
community of students in this department. To all the students who have
since moved on and show what we may be: Bobby, Leo, Ryan, Katie,
Fletcher, Ayden, Maheyer, Jess, Dhwani, Spencer, Mallory, Omar, Isaac,
and Dasha.

And to my friends across UVic, Wyatt, Rosemary, Maggie, and the many
people I am still meeting.

To the Linguistics extended family and team \emph{adequate}. Melody,
Lily, Gillian, Stephen, Jamie, Chloë, Alyssa, Kyle, and Marcus. You are
all wonderful human beings, cooks, and always a joy to spend time with.

To my musical friends, Adrienne, Alex for our shared joy in music
making. And Steve, for showing me the joy in a simple life and beautiful
music.

To the Victoria performing arts community and Fierce Brosnan, our local
legend.

To my fellow crafters, knitters, and crochet friends. Leise, Wes, Anne,
Anya, Ashley, Sabrina, Sarah. To the graduate student society for
continuing to connect us across departments.

To the international community, and my international friends---Marcela,
Bernardo, Daniel, Walla, Mehrab, and everyone else I have met.

To Erica. To my students and everyone working with outreach here. I love
being able to pass on our knowledge to the wider world.

To my TEP friends---I hope you all continue to do well.

To all of my grade school math nerds---Stanley, Alex, Rick, Andrew. I
enjoyed all of our adventures together, appreciate helping myself bloom
in curiosity and exploration, and keep on math-ing!

To my past advisors, James who has continued to support me long after
our paths diverged, David and Jennifer for guiding my first research
project. To Wayne for showing us the joy of astronomy and outreach. To
my mentor, Ebtihal.

To my past roommates, Matthew, Josh, and Bryn, for putting up with my
chaos and our shared time together.

To my long-time friends. Nikola and Eric. Mikey and Anthony. Anya, Kaia,
Alyssa, Aahliyah, Cameron. May we continue to find each other even as we
drift apart.

To my parents for their continuous support throughout this project. To
my grandparents for always being there.

To all of my cousins, but especially Bridget and Gretchen, and my
extended family. And to my siblings for always being a source of joy. To
Arya. To all of my furry friends I have sat---Nano, Pluto, Hobbs,
Penelope, Tochka, Camar, Loki, and Hermes.

To the stars and the moon, for your enduring beauty, continuing to
inspire me and posing unending questions about our natural world.

To the land of Victoria, and its historical caretakers. To the waters
and beautiful trees, the ocean waves, rocks and hills and jagged
coastlines. To the rich, enveloping forests and rust-red arbutus tress,
scraggly Garry oaks and Camus flowers. I could not imagine a more
beautiful place to study and to always remind my place in the world.