### 2025-11-13 thesis/acknowledgements.tex

While written by my hand, I cannot pretend that I alone created this
work. I am ever thankful and greatly indebted to the multitude of people
who have supported, inspired, and nurtured me as a human being.

To my supervisor, Julio Navarro, for being the critical eye and guiding
reason to this work. I will always approach research differently now.

To Raphaël, who has helped inspire and direct this project, and to
everyone who has shared ideas and opportunities along the way: Isabel,
Jens, Eduardo, Jorge, David, Justin, Eugene, Azi, José, and many others.

To Jax for you guidance and personal support for this project.

To all of my new friends I have made along the way---Ruxin, Aviv. To
Pranav, Ted, Jono, Adrianne, Praneet, Will, Maeve, Ewoud, Anne, Anya,
Shoshanah. To Akshara and Simon for all your personal and professional
advice. We have built a wonderful graduate community. To all the
students who have since moved on but still inspire me: Bobby, Leo, Ryan,
Katie, Fletcher, Ayden, Maheyer, Jess, Dhwani, Spencer, Mallory, Omar,
Isaac, and Dasha.

To my friends across UVic: Wyatt, Rosemary, Maggie. To the Linguistics
extended family and team \emph{adequate}: Melody, Lily, Gillian,
Stephen, Jamie, Chloë, Alyssa, Kyle, and Marcus. You are all wonderful
human beings, cooks, and always a joy to spend time with.

To my musical friends, Adrienne and Alex, for our shared joy in music
making. And Steve, for showing me the joy in a simple life and beautiful
music.

To the Victoria performing arts community and Fierce Brosnan, our local
legend.

To my fellow fibre friends and crafters: Leise, Wes, Ashley, Sabrina,
Sarah. To the graduate student society for continuing to connect us
across departments.

To the international community, and my international friends---Marcela,
Bernardo, Daniel, Victor, Mehrab, and many others.

To Erica. To my students and everyone working with outreach here. I love
being able to pass on our knowledge to the wider world.

To my TEP friends---I hope you all continue to do well.

To all of my grade school math nerds---Stanley, Alex, Rick, Andrew. I
enjoyed all of our adventures together, appreciate helping myself bloom
in curiosity and exploration, and keep on math-ing!

To my past advisors, James who has continued to support me long after
our paths diverged, David and Jennifer for guiding my first research
project. To Wayne for showing us the joy of astronomy and outreach. To
my mentor, Ebtihal.

To my past roommates, Matthew, Josh, and Bryn, for putting up with my
chaos and our shared time together.

To my long-time friends---Nikola, Eric, Mikey, Anthony, Anya, Kaia,
Alyssa, Aahliyah, Cameron---may we continue to stay in touch.

To my parents for their continuous support throughout this project. To
my grandparents for always being there. To Nolan and Olivia, always a
joy to see each other. To Bridget and Gretchen. To my extended family.
And to my siblings for always being a source of joy.

To Arya. To all of my furry friends of friends---Nano, Pluto, Hobbs,
Penelope, Tochka, Camar, Loki, and Hermes.

To the stars and the moon, for your enduring beauty, continuing to
inspire me and posing unending questions about our natural world.

To the land of Victoria, and its historical caretakers. To the waters
and beautiful trees, the ocean waves, rocks and hills and jagged
coastlines. To the rich, enveloping forests and rust-red arbutus tress,
scraggly Garry oaks and Camus flowers. I could not imagine a more
beautiful place to study and to always remind my place in the world.